---
title: "epartakki"
author: "Eleni Partakki"
date: "3/17/2021"
output: html_document
---

## Must add changes

## Questions to answer

a. Some activities may deviate from an established pattern, or are just difficult to explain from what you know of a situation. Describe any unusual patterns (either single day or multiple days) and highlight why you find them unusual.
b. What are the top 3 patterns you discovered that you suspect could be most impactful to bird life in the preserve?
c. Turning your attention to the sensor data, characterize the sensors’ performance and operation. - Are they all working properly?
d. Can you detect any unexpected behaviors of the sensors through analyzing the readings they capture?
**existing ideas on tableau and python


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r packages, include=FALSE}
library(tidyverse)
library(readxl)
library(janitor)
```

```{r load-data, message=FALSE}
meteorological_data <- read_xlsx("~/Downloads/DC3-data/Sensor Data/Meteorological Data.xlsx")
sensor_data <- read_xlsx("~/Downloads/DC3-data/Sensor Data/Sensor Data.xlsx")
lekagul_sensor_data <- read_csv("~/Downloads/DC3-data/Traffic Data/Lekagul Sensor Data.csv")
```

```{r data-cleaning}
meteorological_data <- meteorological_data %>%
  clean_names() %>%
  select(-x4)
sensor_data <- sensor_data %>% clean_names()
lekagul_sensor_data <- lekagul_sensor_data %>%
  clean_names() %>%
  mutate(car_type = as.factor(car_type))
```

# Acknowledgements
This dataset was prepared by Pacific Northwest National Laboratory as part of the VAST Challenge 2017.